Deep Learning

### 2015-01-27 deepreview.tex
GEOFF's EDITIONS SENT VIA deepreview_jan26.tex BY EMAIL

Deep neural networks that learn to represent data in multiple layers
of increasing abstraction have dramatically improved the
state-of-the-art for speech recognition, object recognition, object
detection, predicting the activity of drug molecules, and many other
tasks. Deep learning discovers intricate structure in large datasets
by using the backpropagation algorithm to compute error-derivatives
for the weights that define each layer of feature detector and then
repeatedly changing these weights in proportion to their derivatives
averaged over small batches of training examples.  Deep recurrent
neural networks, in which each layer also receives input from its own
state at the previous time-step, can now generate good captions for
images and are competitive for language translation.

@@

Deep learning is currently making major advances in solving problems
that have resisted the best attempts of the Artificial Intelligence
community for many years. Deep learning is very good at discovering
intricate structure in large, high-dimensional datasets and is
therefore applicable to many domains of science and business. For
example, it has beaten previous machine learning techniques at
predicting the activity of potential drug
molecules~\citep{Dahl-et-al-arxiv2014}, detecting new fundamental
particles~\citep{Melis-Higgs-boson-competition-2014}, % TODO ~\citep{YANNparticles}, 
classifying galaxies, % %TODO ~\citep{YANNgalaxies}, 
and predicting the effects of mutations
in non-coding DNA on gene expression and
disease~\citep{Keung-et-al-2014}. % Frey

So far, most applications of deep learning use feed-forward neural
networks, but for tasks that involve sequential inputs, such as speech
and language, it is often better to use recurrent neural
networks. These are given one element from an input 
% YB term-->element 
sequence at each time step and their recurrent internal
``hidden'' state depends on the whole history of the input
sequence. Recurrent networks are very good at predicting the next
character in text~\citep{Sutskever-et-al-ICML2011} or the next
word in a sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be
used for more complex tasks.  After reading an English sentence one
word at a time, for example, an English``encoder'' network can be
trained so that the final state vector of its hidden units is a good
representation of the thought expressed by the sentence.  This thought
vector can then be used as the initial hidden state (or as extra
input) of a trained French ``decoder'' network which outputs a
probability distribution for the first word of the French
translation. If a particular first word is chosen (e.g., according to
that probability) and provided as input to the decoder network it will
then output a probability distribution for the second word of the
translation and so on until a full stop is
chosen~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
Overall, this process generates sequences of French words according to
a probability distribution that depends on the English sentence.

This rather naive way of performing machine translation has quickly become
competitive with the state-of-the-art and this raises serious doubts about
whether understanding a sentence requires anything like internal symbolic
expressions that are manipulated by binding their symbols to the variables
in judiciously chosen rules of inference.  The encoder-decoder translation
networks do not work by using internal rules of inference in which
variables get bound to symbols. They just use big state vectors, big weight
matrices and scalar non-linearities to get the job done.  This suggests
that deriving conclusions using only the form of the premises (which is
what symbolic systems do) is a powerful
technique that people can eventually master but it is not a good model of
the way we normally reason or understand language. Recurrent neural
networks that represent thoughts as big vectors are more compatible with
the view that everyday reasoning involves many simultaneous analogies that
each contribute plausibility to a conclusion~\citep{metaphors}.

This review is organized chronologically starting with the introduction of
the backpropagation algorithm in the 1980's. After describing some of its
early successes, we explain the actual and perceived limitations that
prevented it from living up to its early promise for so many years and how
these limitations were eventually overcome.

@@

Patterns can be classified by first converting each raw input vector into a
vector of feature activations and then collecting evidence for each
possible class by computing a weighted sum of the features. If good
features can be designed by hand, it is relatively easy to learn
class-specific weights that optimize
discrimination~\citep{Rosenblatt57}. Widely-used examples of such linear
classification methods are logistic regression and linear support vector
machines~\citep{bishop-book95,Boser92}.

However, even with good hand-crafted features, many interesting tasks
cannot be solved by linear classifiers. A common method to get around this
problem is to transform the feature vector into a higher dimensional vector
through a non-linear mapping, exploiting the fact that collections of
vectors are more likely to be easily separated when they are spread around
a high-dimensional space~\citep{Cover65}. When there are at least
as many dimensions as examples, a linear classifier can always separate the
training examples, and one clever way to obtain a very high-dimensional space is to use a
``kernel function'' that measures the similarity between two vectors.
Using the kernel function, any input vector, ${\bf x}_i$ in the training
set can be used to define a feature whose activation for a new input
vector, ${\bf x}_{\rm new}$ is just the similarity between ${\bf x}_i$ and
${\bf x}_{\rm new}$. Given a kernel function, it is possible to jointly
optimize the choice of training examplars used to define features and the
class-specific weights associated with the features~\citep{Boser92}. The training
examplars used to define features are called ``support vectors'' and Support
Vector Machines~\citep{Boser92} have been very successful, especially for
tasks with a relatively small number of training examples in which features
are individually informative. They do not, however, perform well for tasks
such as object recognition in which the input space is high-dimensional and
the input-output mapping is so complex that computing similarities with a
simple kernel function is ineffective. With a general-purpose feature space
and kernel (such as the Gaussian kernel), generalization far from the
examples is poor~\citep{Bengio-localfailure-NIPS-2006-small}, which is problematic
in high-dimensional spaces. However, kernel machines would perform well if a
very good set of features were provided: learning very good
features is what deep learning can do.

Problems such as image and speech recognition require the input-output
function to be insensitive to irrelevant variations of the input, such as
variations in position, orientation or illumination of an object, or
variation in pitch or accent of speech, while being very sensitive to
particular minute variations, such as the difference between a white wolf
and a samoyed dog. Experiments and theory of deep learning both
suggest that these tasks are better handled with multiple layers of
progressively more complicated feature detectors each of which is defined
in terms of the simpler feature detectors in the previous layers.

From the earliest days of pattern recognition~\citep{selfridge,Rosenblatt57}
researchers aimed to replace hand-engineered features by multiple layers of
learned features.  This was eventually achieved by using the
backpropagation algorithm which was discovered independently by several
different groups in the 1970s and 1980s~\citep{Werbos74,Parker85,LeCun85,RHW}.  
Backpropagation is simply an application of
the chain rule of derivatives (left of Figure~\ref{fig:backprop-box}),
which yields an efficient way of computing the gradient of a
smooth objective function with respect to all of the weights and parameters in 
a layered network (Figures~\ref{fig:backprop-box} and~\ref{fig:mlp}). 
The objective function typically measures an error to be minimized: 
the discrepancy between
the actual output of the network on a given training case and the target
output for that case.  The key insight is that the error derivatives with
respect to the {\it inputs} of a layer can be computed by working
backwards from the error derivatives with respect to its {\em outputs}. 
Once this backwards recursion has been used to compute
error derivatives for the outputs of all layers, it is easy to compute
error derivatives for the parameters (or weights) of each layer 
(Figure~\ref{fig:mlp}).

%%%%%%% GEOFF %%%%%%%%%%
%GEOFF: Do box explaining backprop equatons
%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.33\linewidth]{chain-rule.pdf}
\includegraphics[width=0.33\linewidth]{forward-graph.pdf}
\includegraphics[width=0.33\linewidth]{backward-graph.pdf}
}
\caption{
Left: the chain rule of derivatives tells us how
two small effects (that of a small change of $x$ on $y$, and that of $y$ on $z$)
are composed. A small change $\Delta x$ in $x$
gets transformed first into a small change $\Delta y$ in $y$
by getting multiplied by $\frac{\partial y}{\partial x}$
(that is the definition of partial derivative). Similarly,
the change $\Delta y$ creates a change $\Delta z$ in $z$. Substituting one equation
into the other gives the chain rule of derivatives, i.e., how
$\Delta x$ gets turned into $\Delta z$ through multiplication
by the product of $\frac{\partial y}{\partial x}$ and
$\frac{\partial z}{\partial x}$. It also works when $x$, $y$
and $z$ are vectors (and the derivatives are Jacobian matrices). \newline
Middle: many computations, such as the succession of linear and non-linear 
transformations of multi-layer neural networks, can be described as
a graph where each node $i$ computes a quantity $u_i$ (e.g. the
output of an artificial neuron or the value of a parameter such as
a synaptic weight), based on the set of values $a_i$, composed of elements $u_j$ of
predecessors in the graph, i.e., $u_i = f_i(a_i)$. This can be readily
generalized to nodes whose output is a vector or a matrix.\newline
Right: the chain rule can be applied efficiently to computational
flow graphs by recursively computing the derivative of the final
node $u_N$ (e.g., the error to be minimized or the output of the neural network)
with respect to every internal node. The basic recursive equation
is
$\frac{\partial u_N}{\partial u_j} = \sum_{i: j \in a_i} \frac{\partial u_N}{\partial u_i}
\frac{\partial f(a_i)}{\partial u_j}.$
}
\label{fig:backprop-box}
\end{minipage}
}
\end{figure}

@@

Backpropagation provides an efficient way of computing gradients with respect to all of
the weights in a layered network on a single training case, but there are
still many ways to make use of these gradients. The most obvious method is
to average the gradients over all of the cases in the training set and then
to change each weight in proportion to its negative gradient, which is the
direction where a small change will most decrease the objective function. 
When the training set contains millions of cases and is highly redundant
% explaining what ``redundant'' means, here
(with many examples illustrating variations of the same underlying 
aspect or concept), however,
this is needlessly slow. It is much faster to compute the average gradient
on a small ``mini-batch'' of randomly chosen training cases and to update
each weight by this stochastic estimate of its full gradient, times a
learning rate which decays as learning proceeds. Stochastic gradient
descent can be improved by using the history of the individual weight
gradients to adapt the learning rate for each weight~\citep{bottou-bousquet-2008-small}
or to provide additional increments that depend on earlier gradients 
but these techniques are still the subject of controversy and are beyond
the scope of this review (but see~\citet{Montavon2012} for tricks of the trade). 
However, it is now widely accepted that some
form of stochastic gradient descent is the best bet for large, redundant
datasets.

@@

The connectivity of the network is restricted so that each of the one-of-N
input vectors must first be converted to a ``word-vector'' before being
used by subsequent layers to give a high probability to the target next
word.  The network learns word-vectors that contain many active components
each of which can be interpreted as a separate feature of the word. This
was first demonstrated using sequences of three words, such as
(Colin, Mother, Victoria), derived from two family trees~\citep{RHW}. The learned word
vector for Colin contained ``semantic'' features representing which family
tree Colin was in, what branch of that tree he was in, and from what generation
he was. These semantic features were not explicitly present
in the input.  They were discovered by the learning procedure as a good way
of factorizing the structured relationships between the input and output
symbols into multiple ``micro-rules''.  For example, given Colin's
generation and the learned feature of Mother that the output should be one
generation higher than the input, we can already restrict the answer to
people in a specific generation. The ability of deep learning to discover
many such micro-rules allows it to generalize well to novel combinations of
individually familiar context words.

@@

The issue of representation lies at the heart of the debate between the
logic-inspired and the neurally-inspired paradigms for cognition. In the
logic-inspired paradigm, an instance of a symbol is something whose only
property is that it is either identical or non-identical to other symbol
instances. It has no internal structure that is relevant to its use. By
contrast, neural networks use large activity vectors instead of symbols and
parallel matrix operations followed by non-linearities instead of formal
rules of inference. The elements of such vectors can be understood as
the values of learned attributes (or features) associated with some object. 
With sufficient data, computational power and depth,
the combination of backpropagation for computing gradients and stochastic
gradient descent for updating the weights on connections has proved to be
very good at learning appropriate distributed representations. These
representations allow a lot of parallel computation at the feature level so
that complex tasks can be solved in a very small number of sequential
steps. In fact, it can be shown that a measure of the richness of the functions
that can be represented by a deep neural network grows exponentially
with depth~\citep{Montufar-et-al-NIPS2014}, whereas according to the
same measure, classical (shallow) non-parametric statistical methods such as
Gaussian Support Vector Machines~\citep{Bengio-localfailure-NIPS-2006-small} 
or decision trees~\citep{Bengio-decision-trees10}
would require an exponentially large number
of training examples to represent the same kinds of functions.

@@

%%%%%%%% YANN %%%%%%% 
% YANN: can you summarize convnets and their use for reading cheques in about
% 350 words?
% I couldn't do it in 350 words.

@@

One of the most widely used deep learning architectures is the
convolutional networt, or ConvNet~\cite{lecun-90c,lecun-98}. ConvNets
are designed to process data that come in the form of arrays, for
example a color image composed of three 2D arrays containing pixel
intensities in the three color channels. Many data modalities are in
the form of multiple arrays: 1D for signals and language, 2D for
images or audio spectrograms, 3D for video or volumetric images.
There are four key ideas behind ConvNets that take advantage of the
properties of many such natural signals: local connections, shared
weights, pooling, and the use of many layers.

The architecture of a typical ConvNet is 
%shown in
%figure~\ref{fig-convnet}. 
structured as a series of stages. The first few stages are composed of two
types of layers: convolutional layers and pooling layers. Units in a
convolutional layer are organized in {\em faeture maps}, within which
each unit is connected to local patches in the feature maps of the
previous layer through a set of weights called a {\em filter
  bank}. The result of this local weight sum is then passed through a
non-linearity.  All units in a feature map share the same filter
bank. Multiple feature maps in a layer use different filter banks. The
reason for this architecture is twofold. First, In array data such as
images, local groups of values are often highly correlated forming
local motifs that are easily detected. Second, the local statistics of
images and other signals are invariant to location. In other word, if
a motif can appear in one part of the image, it could appear anywhere,
hence the idea of units at different locations sharing the same
weights.  Mathematically, the filtering operation performed by a
feature map is a {\em discrete convolution}, hence the name
ConvNet. While the role of the convolutional layer is to detect local
conjunctions of features in the previous layer, the role of the
pooling layer is to merge semantically similar features into
one. Since the relative positions of the features forming a motif can
vary somewhat, reliably detecting the motif can be done by
coarse-graining the position of each feature. A typical pooling unit
computes the max of a local patch of units in one feature map (or in a
few). Neighboring pooling units take input from patches that are
shifted by more than one row or column, thereby reducing the dimension
of the representation and creating an invariance to small shifts.  Two
or three stages of convolutionas, non-linearity, and pooling are
stacked, followed by more convolutional and fully-connected layers.
Back-propagating gradients through a ConvNet is as simple
as through a regular deep network, allowing all the weights in all the
filter banks to be trained.

This architecture exploits the property that many natural signals are
{\em compositional hierarchies}. In images, local combinations of
edges form motifs, motifs assemble into parts, and parts form objects.
Similar hierarchies exists in speech and text from sounds to phones,
phonemes, syllables, words, and sentences. The elements that form
elements in the next level can vary in relative position and
appearance. The convolutional and pooling layers in ConvNets are
directly inspired by the classic notions of simple cells and complex
cells in visual neuroscience~\cite{Hubel62}, and the overall
architecture is reminiscent of visual cortex ventral pathway hierarchy
LGN-V1-V2-V4-IT~\cite{Felleman+VanEssen-1991}. ConvNets have their roots in
Fukushima's Neocognitron~\cite{fukushima-82} whose architecture was
somewhat similar but did not have an end-to-end supervised learning
algorithm such as backpropagation+SGD.

There have been numerous applications of convolutional networks going
back to the early 1990s. Perhaps the earliest success was a bank check
reading system which was commercially deployed by AT\&T and NCR in
1994~\cite{lecun-98}. The system used a ConvNets trained jointly with
a graphical model that implemented language constraints. By the late
1990s this system was reading over 10\% of all the checks in the
US. Microsoft later deploy ConvNets in a number of OCR and handwriting
recognition
systems~\cite{simard-03,chellapilla-ist-06,chellapilla-iwfhr-06b}
including for Arabic and
Chinese~\cite{abdulkader-iwfhr-06,chellapilla-iwfhr-06a}. ConvNets
have also been used since the early 1990s for object detection in
natural images, including faces and
pedestrians~\cite{vaillant-monrocq-lecun-94,garcia-delakis-04,osadchy-07,nasse-09,sermanet-cvpr-13}.

@@

Recurrent neural networks are very powerful dynamical systems, but training
them with BPTT proved to be problematic, because the backpropagated
gradients either grow or shrink at each time step, so over many time-steps
they typically explode or vanish~\citep{Hochreiter91-small,Bengio-et-al-TNN1994},
and the conditions which allow stable dynamics also correspond to
vanishing gradients~\citep{Bengio-et-al-TNN1994},

@@

The obvious weakness of gradient descent is that it can get trapped in poor
local minima, but in high-dimensional weight-spaces this is much less
common than getting stuck close to saddle points where the error surface
curves upwards in most directions but downwards in a few. Recent
results~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015} strongly suggest
the vast majority of critical points (where derivatives of the total
objective function with respect to parameters vanish) are saddle points, not
local minima, and that in large deep networks almost all of these local minima 
give very similar performance so it does not matter which one we find.  This goes a long way
towards explaining why stochastic gradient descent works so well for deep
neural networks.

>>>>

Deep neural networks that learn to represent data using layers
of increasing abstract features have dramatically improved the
state-of-the-art for speech recognition, object recognition, object
detection, predicting the activity of drug molecules, and many other
tasks. Deep learning discovers intricate structure in large datasets
by using the backpropagation algorithm to compute error-derivatives
for the weights that define each layer of feature detectors and then
repeatedly changing these weights in proportion to their derivatives
averaged over small batches of training examples.  Deep recurrent
neural networks, in which each layer also receives input from its own
state at the previous time-step, can now generate good captions for
images and are competitive for language translation.

@@

Deep learning is making major advances in solving problems
that have resisted the best attempts of the Artificial Intelligence
community for many years. Deep learning is very good at discovering
intricate structure in large, high-dimensional datasets and is
therefore applicable to many domains of science and business. For
example, it has beaten previous machine learning techniques at
predicting the activity of potential drug
molecules~\citep{Dahl-et-al-arxiv2014}, detecting new fundamental
particles~\citep{Melis-Higgs-boson-competition-2014}, % TODO ~\citep{YANNparticles}, 
classifying galaxies, % %TODO ~\citep{YANNgalaxies}, 
and predicting the effects of mutations
in non-coding DNA on gene expression and
disease~\citep{Keung-et-al-2014}. % Frey

So far, most applications of deep learning use feed-forward neural
networks, but for tasks that involve sequential inputs, such as speech
and language, it is often better to use recurrent neural
networks. These are given one element from an input 
% YB term-->element 
sequence at each time step and their recurrent internal
``hidden'' state depends on the whole history of the input
sequence. Recurrent networks are very good at predicting the next
character in text~\citep{Sutskever-et-al-ICML2011} or the next
word in a sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be
used for more complex tasks.  After reading an English sentence one
word at a time, for example, an English ``encoder'' network can be
trained so that the final state vector of its hidden units is a good
representation of the thought expressed by the sentence.  This thought
vector can then be used as the initial hidden state of (or as extra
input to)  a jointly trained French ``decoder'' network which outputs a
probability distribution for the first word of the French
translation. If a particular first word is chosen from this distribution and provided as input to the decoder network it will
then output a probability distribution for the second word of the
translation and so on until a full stop is
chosen~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
Overall, this process generates sequences of French words according to
a probability distribution that depends on the English sentence.

This rather naive way of performing machine translation has quickly become
competitive with the state-of-the-art and this raises serious doubts about
whether understanding a sentence requires anything like internal symbolic
expressions that are manipulated by binding their symbols to the variables
in judiciously chosen rules of inference.  The encoder-decoder translation
networks do not work by using internal rules of inference in which
variables get bound to symbols. They just use big state vectors, big weight
matrices and scalar non-linearities to get the job done.  Deriving conclusions using only the form of the premises (which is
what formal logic does) is a powerful
technique that people can eventually master, but the translation
results with recurrent nets suggest that it may not be a good model of
the way we normally reason or understand language. Recurrent neural
networks that represent thoughts as big vectors are more compatible with
the view that everyday reasoning involves many simultaneous analogies that
each contribute plausibility to a conclusion~\citep{metaphors,rogers}.

This review is organized chronologically. We start with the introduction of
the backpropagation algorithm in the 1980's and early demonstrations
that it could learn good features for language and vision tasks.  We
then explain the actual and perceived limitations that
prevented it from living up to its early promise for so many years and how
these limitations were eventually overcome.

@@

Patterns can be classified by first converting each raw input vector into a
vector of feature activations and then collecting evidence for each
different class by computing a differently weighted sum of the feature
activations. If good
features can be designed by hand, it is relatively easy to learn
class-specific weights that optimize
discrimination~\citep{Rosenblatt57}. Widely-used examples of such linear
classification methods are logistic regression and linear support vector
machines~\citep{bishop-book95,Boser92}.

However, even with good hand-crafted features, many interesting tasks
cannot be solved by linear classifiers. A common method to get around this
problem is to transform the feature vector into a higher dimensional vector
through a non-linear mapping, exploiting the fact that collections of
vectors are more likely to be easily separated when they are spread around
a high-dimensional space~\citep{Cover65}. When there are at least
as many dimensions as examples, a linear classifier can always separate the
training examples, and one clever way to obtain a very high-dimensional space is to use a
``kernel function'' that measures the similarity between two vectors.
Using the kernel function, any input vector, ${\bf x}_i$ in the training
set can be used to define a feature whose activation for a new input
vector, ${\bf x}_{\rm new}$ is just the similarity between ${\bf x}_i$ and
${\bf x}_{\rm new}$. Given a kernel function, it is possible to jointly
optimize the choice of training examplars used to define features and the
class-specific weights associated with the features~\citep{Boser92}. The training
examplars used to define features are called ``support vectors'' and Support
Vector Machines~\citep{Boser92} have been very successful, especially for
tasks with a relatively small number of training examples in which features
are individually informative. They do not, however, perform well for tasks
such as object recognition in which the input space is high-dimensional and
the input-output mapping is so complex that computing similarities using a
simple kernel function is ineffective. 
%% THE TEXT BELOW IS NICE BUT TOO MUCH DETAIL ABOUT  SVMS FOR THIS
%% REVIEW.  
%% With a general-purpose feature space
%% and kernel (such as the Gaussian kernel), generalization far from the
%% examples is poor~\citep{Bengio-localfailure-NIPS-2006-small}, which is problematic
%% in high-dimensional spaces. However, kernel machines would perform well if a
%% very good set of features were provided: learning very good
%% features is what deep learning can do.

Problems such as image and speech recognition require the input-output
function to be insensitive to irrelevant variations of the input, such as
variations in position, orientation or illumination of an object, or
variation in pitch or accent of speech, while being very sensitive to
particular minute variations, such as the difference between a white wolf
and a samoyed dog. Experimental and theoretical deep learning results 
suggest that these tasks are better handled with multiple layers of
progressively more complicated feature detectors each of which is defined
in terms of the simpler feature detectors in the previous layers.

From the earliest days of pattern recognition~\citep{selfridge,Rosenblatt57}
researchers aimed to replace hand-engineered features by multiple layers of
learned features.  This was eventually achieved by using the
backpropagation algorithm which was discovered independently by several
different groups in the 1970s and 1980s~\citep{Werbos74,Parker85,LeCun85,RHW}.  
Backpropagation is simply an application of
the chain rule of derivatives (left of Figure~\ref{fig:backprop-box}),
which yields an efficient way of computing the gradient of a
smooth objective function with respect to all of the weights 
%% and parameters 
%% FOR THE NOVICE READER INCLUDING BOTH WEIGHTS AND "PARAMTERS" IS CONFUSING.
%% ITS BETTER TO SAY SOMETHING CLEARER EVEN IF IT IS LESS GENERAL OR CORRECT.
in  a layered network (Figures~\ref{fig:backprop-box} and~\ref{fig:mlp}). 
The objective function typically measures an error to be minimized: 
the discrepancy between
the actual output of the network on a given training case and the target
output for that case.  The key insight is that the error derivatives with
respect to the {\it ouputs} of a layer can be computed by working
backwards from the error derivatives with respect to its {\em inputs}
to the subsequent layer.
%% THE TRICKY STEP IS COMPUTING dE/dOUT FROM dE/dIN IN THE SUBSEQUENT
%% LAYER.  I THINK WE SHOULD FOCUS ON GETTING OVER THAT IDEA FIRST.
%%  I PLAN TO REDO THE FIGURE TO MAKE IT SIMPLER.  ITS MUCH TOO
%%  CONFUSING FOR A NOVICE AT PRESENT.  WE SHOULD HAVE A SEPARATE
%%  DISCUSSION OF THAT.
Once this backwards recursion has been used to compute
error derivatives for the outputs of all layers, it is easy to compute
error derivatives for the incoming weights of the neurons in each layer 
(Figure~\ref{fig:mlp}).

\begin{figure}[htp]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.33\linewidth]{chain-rule.pdf}
\includegraphics[width=0.33\linewidth]{forward-graph.pdf}
\includegraphics[width=0.33\linewidth]{backward-graph.pdf}
}
\caption{
Left: the chain rule of derivatives tells us how
two small effects (that of a small change of $x$ on $y$, and that of $y$ on $z$)
are composed. A small change $\Delta x$ in $x$
gets transformed first into a small change $\Delta y$ in $y$
by getting multiplied by $\frac{\partial y}{\partial x}$
(that is the definition of partial derivative). Similarly,
the change $\Delta y$ creates a change $\Delta z$ in $z$. Substituting one equation
into the other gives the chain rule of derivatives, i.e., how
$\Delta x$ gets turned into $\Delta z$ through multiplication
by the product of $\frac{\partial y}{\partial x}$ and
$\frac{\partial z}{\partial x}$. It also works when $x$, $y$
and $z$ are vectors (and the derivatives are Jacobian matrices). \newline
Middle: many computations, such as the succession of linear and non-linear 
transformations of multi-layer neural networks, can be described as
a graph where each node $i$ computes a quantity $u_i$ (e.g. the
output of an artificial neuron or the value of a parameter such as
a synaptic weight), based on the set of values $a_i$, composed of elements $u_j$ of
predecessors in the graph, i.e., $u_i = f_i(a_i)$. This can be readily
generalized to nodes whose output is a vector or a matrix.\newline
Right: the chain rule can be applied efficiently to computational
flow graphs by recursively computing the derivative of the final
node $u_N$ (e.g., the error to be minimized or the output of the neural network)
with respect to every internal node. The basic recursive equation
is
$\frac{\partial u_N}{\partial u_j} = \sum_{i: j \in a_i} \frac{\partial u_N}{\partial u_i}
\frac{\partial f(a_i)}{\partial u_j}.$
}
\label{fig:backprop-box}
\end{minipage}
}
\end{figure}

@@

Backpropagation provides an efficient way of computing gradients with respect to all of
the weights in a layered network on a single training case, but there are
still many ways to make use of these gradients. The most obvious method is
to average the gradients over all of the cases in the training set and then
to change each weight in proportion to its negative gradient. This
moves the weight vector in the direction of steepest descent in the
objective function. However, this is needlessly slow when the 
training set contains millions of cases and is highly redundant
with most examples conveying no new information about the mapping from
inputs to outputs.  It is much faster to compute the average gradient
on a small ``mini-batch'' of randomly chosen training cases and to update
each weight by this stochastic estimate of its full gradient, times a
learning rate which decays as learning proceeds. Stochastic gradient
descent can be improved by using the history of the individual weight
gradients to adapt the learning rate for each weight~\citep{bottou-bousquet-2008-small}
or to provide additional increments that depend on earlier gradients 
but these techniques are still the subject of controversy and are beyond
the scope of this review (see~\citet{Montavon2012} for tricks of the trade). 
However, it is now widely accepted that some
form of stochastic gradient descent is the best bet for large, redundant
datasets.

@@

The connectivity of the network is restricted so that each of the one-of-N
input vectors must first be converted to a ``word-vector'' before being
used by subsequent layers to give a high probability to the target next
word.  The network learns word-vectors that contain many active components
each of which can be interpreted as a separate feature of the word. This
was first demonstrated using sequences of three words, such as
(Colin, Mother, Victoria), derived from two family trees~\citep{RHW}. The learned word
vector for Colin contained ``semantic'' features representing which family
tree Colin was in, what branch of that tree he was in, and what generation
he was from. 
%% TWO NATIVE ENGLISH SPEAKERS AGREE THAT THE "FROM" SHOULD STAY AT THE
%% END IN ORDER NOT TO SOUND STILTED.
These semantic features were not explicitly present
in the input.  They were discovered by the learning procedure as a good way
of factorizing the structured relationships between the input and output
symbols into multiple ``micro-rules''.  For example, given Colin's
generation and the learned feature of Mother that the output should be one
generation higher than the input, we can already restrict the answer to
people in a specific generation. The ability of deep learning to discover
many such micro-rules allows it to generalize well to novel combinations of
individually familiar context words.

@@

The issue of representation lies at the heart of the debate between the
logic-inspired and the neurally-inspired paradigms for cognition. In the
logic-inspired paradigm, an instance of a symbol is something whose only
property is that it is either identical or non-identical to other symbol
instances. It has no internal structure that is relevant to its use. By
contrast, neural networks use large activity vectors instead of symbols and
parallel matrix operations followed by non-linearities instead of formal
rules of inference. The elements of such vectors can be understood as
the values of learned attributes (or features) associated with some object. 
With sufficient data, computational power and depth,
the combination of backpropagation for computing gradients and stochastic
gradient descent for updating the weights on connections has proved to be
very good at learning appropriate distributed representations. These
representations allow a lot of parallel computation at the feature level so
that complex tasks can be solved in a very small number of sequential
steps. In fact, it can be shown that a measure of the richness of the functions
that can be represented by a deep neural network grows exponentially
with depth~\citep{Montufar-et-al-NIPS2014}, whereas according to the
same measure, classical (shallow) non-parametric statistical methods such as
Gaussian Support Vector Machines~\citep{Bengio-localfailure-NIPS-2006-small} 
or decision trees~\citep{Bengio-decision-trees10}
would require exponential storage 
%% an exponentially large number of training examples 
%% THIS WOULD BE VERY CONFUSING FOR MOST READERS WHO WILL NOT KNOW
%% ABOUT NON-PARAMETERIC METHODS.
to represent the same kinds of functions.

@@

@@

One of the most widely used deep learning architectures is the
convolutional neural network, or ConvNet~\cite{lecun-90c,lecun-98}. ConvNets
are designed to process data that come in the form of arrays, for
example a color image composed of three 2D arrays containing pixel
intensities in the three color channels. Many data modalities are in
the form of multiple arrays: 1D for signals and language, 2D for
images or audio spectrograms, 3D for video or volumetric images.
There are four key ideas behind ConvNets that take advantage of the
properties of many such natural signals: local connections, shared
weights, pooling, and the use of many layers.

The architecture of a typical ConvNet is 
%shown in
%figure~\ref{fig-convnet}. 
structured as a series of stages. The first few stages are composed of two
types of layers: convolutional layers and pooling layers. Units in a
convolutional layer are organized in {\em feature maps}, within which
each unit is connected to local patches in the feature maps of the
previous layer through a set of weights called a {\em filter
  bank}. The result of this local weight sum is then passed through a
non-linearity.  All units in a feature map share the same filter
bank. Different feature maps in a layer use different filter banks. The
reason for this architecture is twofold. First, in array data such as
images, local groups of values are often highly correlated forming
local motifs that are easily detected. Second, the local statistics of
images and other signals are invariant to location. In other word, if
a motif can appear in one part of the image, it could appear anywhere,
hence the idea of units at different locations sharing the same
weights.  Mathematically, the filtering operation performed by a
feature map is a {\em discrete convolution}, hence the name
ConvNet.

While the role of the convolutional layer is to detect local
conjunctions of features in the previous layer, the role of the
pooling layer is to merge semantically similar features into
one. Since the relative positions of the features forming a motif can
vary somewhat, reliably detecting the motif can be done by
coarse-graining the position of each feature. A typical pooling unit
computes the max of a local patch of units in one feature map (or in a
few). Neighboring pooling units take input from patches that are
shifted by more than one row or column, thereby reducing the dimension
of the representation and creating an invariance to small shifts.  Two
or three stages of convolution, non-linearity, and pooling are
stacked, followed by more convolutional and fully-connected layers.
Back-propagating gradients through a ConvNet is as simple
as through a regular deep network, allowing all the weights in all the
filter banks to be trained. The error derivative for a shared weight
is just the sum the derivatives for all the instances of that weight.

This architecture exploits the property that many natural signals are
{\em compositional hierarchies}. In images, local combinations of
edges form motifs, motifs assemble into parts, and parts form objects.
Similar hierarchies exists in speech and text from sounds to phones,
phonemes, syllables, words, and sentences. The features that form
elements at the next level can vary in relative position and
appearance. The convolutional and pooling layers in ConvNets are
directly inspired by the classic notions of simple cells and complex
cells in visual neuroscience~\cite{Hubel62}, and the overall
architecture is reminiscent of visual cortex ventral pathway hierarchy
LGN-V1-V2-V4-IT~\citep{Felleman+VanEssen-1991}. ConvNets have their roots in
Fukushima's Neocognitron~\citep{fukushima-82} whose architecture was
somewhat similar but did not have an end-to-end supervised learning
algorithm such as backpropagation+SGD.

There have been numerous applications of convolutional networks going
back to the early 1990s. A major early success
%% THIS REWORDING AVOIDS COMPLAINTS FROM HECHT NIELSEN ETC.
was a bank check
reading system which was commercially deployed by AT\&T and NCR in
1994~\cite{lecun-98}. The system used a ConvNets trained jointly with
a graphical model that implemented language constraints. By the late
1990s this system was reading over 10\% of all the checks in the
US. Microsoft later deploy ConvNets in a number of OCR and handwriting
recognition
systems~\cite{simard-03,chellapilla-ist-06,chellapilla-iwfhr-06b}
including for Arabic and
Chinese~\cite{abdulkader-iwfhr-06,chellapilla-iwfhr-06a}. ConvNets
have also been used since the early 1990s for object detection in
natural images, including faces and
pedestrians~\cite{vaillant-monrocq-lecun-94,garcia-delakis-04,osadchy-07,nasse-09,sermanet-cvpr-13}.

@@

Recurrent neural networks are very powerful dynamical systems, but training
them with BPTT proved to be problematic, because the backpropagated
gradients either grow or shrink at each time step, so over many time-steps
they typically explode or
vanish~\citep{Hochreiter91-small,Bengio-et-al-TNN1994}.
%% THE TEXT BELOW IS TOO COMPLICATED AND ALSO TOO SIMPLE: THERE ARE
%% STABLE CONDITIONS IN WHICH THE GRADIENTS DO NOT VANISH (ORTHONORMAL
%% RECURRENT MATRICES). I THINK ITS BETTER NOT TO GET INTO THIS.
%% and the conditions which allow stable dynamics also correspond to
%% vanishing gradients~\citep{Bengio-et-al-TNN1994},

@@

The obvious weakness of gradient descent is that it can get trapped in poor
local minima, but in high-dimensional weight-spaces this is much less
common than getting stuck close to saddle points where the error surface
curves upwards in most directions but downwards in a few. Recent
results~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015} strongly suggest
the vast majority of the critical points (where derivatives of the total
objective function with respect to the parameters vanish) are saddle points, not
local minima, and that in large, deep networks almost all of the 
saddle points with predominantly upwards curvature  
give very similar performance so it does not matter which one we find.  This goes a long way
towards explaining why stochastic gradient descent works so well for deep
neural networks.
